\Large \bf New Static Analysis Techniques to Detect Entropy Failure Vulnerabilities

### 2018-12-12 paper/paper.tex
presentation and paper start

%make title bold and 14 pt font (Latex default is non-bold, 16 pt)
\title{\Large \bf CS380S: Project Update}

@@

One common misuse of cryptography is the misuse of entropy. Without proper random inputs, many cryptographic algorithms
are vulnerable to basic forms of cryptanalysis. Some cryptographic schemes, such as DSA, can even disclose long-term secrets
like the signing key when the random per-message input is low entropy, made public, or nonunique.

We plan to use data dependency tools to determine how entropic inputs in a given program are used by various cryptographic algorithms. That will allow us to identify if and when entropy is too low or is misused. We plan to either produce this as a code integration tool for developers to use as part of a compiler toolchain, or use this tool to analyze a large number of codebases found ``in the wild,'' such as those written by amateurs.
Our tool will seek to generate a human-checkable dependency graph from source code using notions of taint analysis.

@@

@@

One technique that we have considered is the notion of a ``product program.'' In the most general sense, the product $P_1 \times P_2$ of two programs $P_1$ and $P_2$ is used to verify
relations between the programs, such as equivalence. Product programs have also been used to analyze different runs of the same program. 
The product program $P_1 \times P_2$ is semantically equivalent to the sequential composition $P_1; P_2$, but such that we can prove useful safety properties of $P_1 \times P_2$ that would be difficult to prove with standard techniques on $P_1$, $P_2$, or $P_1; P_2$.

In our setting, we are not so much interested in proving safety properties of a program $P$ but instead wish to prove that cryptographic values rely on sufficiently many bits of entropy when they are used via taint analysis.
The notion of product programs helps us introduces the concept of ``differential taint analysis'' which will give u

Given two versions $P_1, P_2$ of the same program P, we propose a relational verification for the entropy of the cryptographic values in the program. By  
we hope to produce a The safety property we are the most interested in proving is that

\section{Example}

@@

\section{Existing Techniques, DTA Motivation}

Static code analysis has a history of identifying security vulnerabilities at a source code level. Some examples include SQL injection, cross-site scripting exploits, and buffer overflow attacks. However, there has not been any attempt in the literature to statically analyze source code for cryptographic vulnerabilities stemming from entropy misuse, which we seek to do. We claim that standard static code analysis techniques developed thus far are insufficient for the analysis we would like to perform.

The sources of entropy can be tainted standard taint analysis techniques can be used to approximate how the taint propagates. But this approximation may be too coarse to provide useful information to the user. Even with a sufficiently precise taint analysis, it is unclear how to determine the ``correct'' set of sources a cryptographic value should rely on at any point in the program.

Verification modulo versions can be leveraged, the static analyzer is treated as opaque in an effort to cut down the number of alarms by inferring assumptions made by the original program (assumed to be correct). This is especially appealing in the security software development cycle because there are documented cases of entropy bugs being introduced in new versions of programs (for example, OpenSSL and FreeBSD). This would involve under-approximating the taint of one version of the program and over-approximating the taint of another version of the program. Current tooling does not widely support under-approximating taint analysis.

Another existing technique is relational verification, in which relational properties between two programs are used to prove $k$-safety properties. In our case, we would want to show the equivalence of entropy assignment and the equivalence of the entropy propagation. Relational verification would proceed with the construction of a product program between versions of the program that captures the semantics of executing them sequentially, but that lends itself to standard verification techniques.

@@

How our analysis framework would be able to identify real world bugs. What are the benchmarks we compare against (OpenSSL, FreeBSD, etc.). What are the projects we will evaluate the tool on. %TODO

@@

\begin{enumerate}
	\item An automated tool can detect entropy bugs in real-world programs.
	\item Entropy is insufficiently propagated in programs that rely on cryptography.
	\item Multiple versions of the same program can make static analysis for this domain more effective
\end{enumerate}

>>>>

%make title bold and 14 pt font (Latex default is non-bold, 16 pt)
\title{\Large \bf New Static Analysis Techniques to Detect Entropy Failure Vulnerabilities}

@@

One common misuse of cryptography is the misuse of entropy. Without proper random inputs, many cryptographic algorithms
are vulnerable to basic forms of cryptanalysis. Some cryptographic schemes, such as DSA, can even disclose long-term secrets
such as the signing key when the random per-message input is low entropy, made public, or nonunique.

We plan to use data dependency tools to determine how entropic inputs in a given program are used by various cryptographic algorithms. That will allow us to identify if and when entropy is too low or is misused. We plan to produce this as a code integration tool for developers to use as part of a compiler toolchain.
Our tool will seek to generate an error report from source code using notions of taint analysis.

@@

\section{Preliminaries}

\subsection{Taint Analysis}

\subsection{Predicate Abstraction}

\subsection{$k$-safety properties}

\subsection{Product Programs}

\section{Our Approach}

\subsection{Instrumenting Programs for Predicate Abstraction}

\subsection{Naive Product Program Inference Rules}

\subsection{Taint-augmented Inference Rules}

\subsection{Putting it all together}

\section{Results}

\section{Conclusion}

\section{Future Work}

@@

\subsection{Taint Analysis}
Static code analysis has a history of identifying security vulnerabilities at a source code level. Some examples include SQL injection, cross-site scripting exploits, and buffer overflow attacks. However, there has not been any attempt in the literature to statically analyze source code for cryptographic vulnerabilities stemming from entropy misuse, which we seek to do. We claim that standard static code analysis techniques developed thus far are insufficient for the analysis we would like to perform. 
The sources of entropy can be tainted standard taint analysis techniques can be used to approximate how the taint propagates. But this approximation may be too coarse to provide useful information to the user. Even with a sufficiently precise taint analysis, it is unclear how to determine the ``correct'' set of sources a cryptographic value should rely on at any point in the program.

\subsection{Differential Taint Analysis}

Since we want a stronger notion of taint analysis than a naive approach, we intend to introduce and formalize the concept of ``differential taint analysis." Differential
taint analysis, in the ideal, will give us more information than simple taint analysis by leveraging 
the original version of the program as a source of ground truth. This additional information could include reducing programmer annotation work or reduce false positives between versions
(since taint analysis overapproximates data flow). By viewing the original version of the program as bug-free, we reduce the number of code audits required:
only the original version of the code would need to be audited, and future versions could rely on our tool to automatically detect bugs that were introduced by refactors.

\subsection{Product Programs}

One technique that we have considered toward the construction of differential taint analysis is the notion of a ``product program.'' In the most general sense, the product $P_1 \times P_2$ of two programs $P_1$ and $P_2$ is used to verify
relations between the programs, such as equivalence. Product programs have also been used to analyze different runs of the same program. 
The product program $P_1 \times P_2$ is semantically equivalent to the sequential composition $P_1; P_2$, but such that we can prove useful safety properties of $P_1 \times P_2$ that would be difficult to prove with standard techniques on $P_1$, $P_2$, or $P_1; P_2$.

In our setting, we are not so much interested in proving safety properties of a program $P$ but instead wish to prove that cryptographic values rely on sufficiently many bits of entropy when they are used via taint analysis, 
which is why we cannot simply leverage existing off-the-shelf product program analysis techniques.

\subsubsection{Example}

@@

\subsection{Verification Modulo Versions}
Verification modulo versions is an idea quite similar to product programs. VMV attempts to identify abstract regressions and relative correctness between two versions of a program.
In VMV, the static analyzer is treated as opaque in an effort to cut down the number of alarms by inferring assumptions made by the original program (assumed to be correct). This is especially appealing in the security software development cycle because there are documented cases of entropy bugs being introduced in new versions of programs (for example, OpenSSL and FreeBSD). This would involve under-approximating the taint of one version of the program and over-approximating the taint of another version of the program. Current tooling does not widely support under-approximating taint analysis.

\subsection{Relational Verification}
Another existing technique is relational verification, in which relational properties between two programs are used to prove $k$-safety properties. A $k$-safety property is a safety property of a program that requires 
reasoning about the relationships between $k$ different runs of a program (or $k$ programs). For example, transitivity is a $k$-safety property.
In our case, we would want to show the equivalence of entropy assignment and the equivalence of the entropy propagation. Relational verification utilitzes product programs; RV constructs a product program between versions of the program that captures the semantics of executing them sequentially, but that lends itself to standard verification techniques.

\subsection{Differential Assertion Checking}

The last approach that we have considered is differential assertion checking. Differential assertion checking seeks to (statically) prove the relative correctness between two similar programs,
with a significantly lower cost than ensuring absolute correctness. That is, DAC allows is to answer if there are conditions under which a program $P_1$ passes an assertion check but
a program $P_2$ fails. The techniques utilized current DAC work may be of interest, but it seems they rely on an a priori mapping between the two programs, which is something that
relational verification work seeks to answer. Thus DAC will be considered further after the idea of product programs is exhausted.

@@

Our tool should be able to detect the infamous Debian OpenSSL entropy bug. We will also test our tool on other libraries that rely on entropy, such as Amazon's signal2noise, LibreSSL, GnuTLS, and GnuPG,
which are not known to contain entropy bugs (and likely do not), but we can test on versions in which we introduce our own bugs.

@@

\begin{enumerate}
	\item An automated tool can detect entropy bugs in real-world programs.
	\item Entropy is insufficiently propagated in programs that rely on cryptography, or entropy propagation follows nontrivial code paths (due to error handling or other control flow).
	\item Multiple versions of the same program can make static analysis for this domain more effective by lowering costs (computational and programmer) or providing more fine-grained information.
\end{enumerate}

\section{Future Work}

As bad data flow can be a source of security bugs other than simply (lack of) entropy propagation (for example, 
address disclosure and use-after-free bugs in browsers), we believe that our tool could also be applied to these settings to help prevent
programmers from introducing security vulnerabilities into their projects.
